---
title: "Movement Classification"
author: "Niyazi Ülke"
date: "06/10/2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview

This is a course project from Coursera Reproducible Research MOOC by JHU. In this project, data from accelometers on the belt, forearm, arm and dumbell of 6 participants will be used. Source of data is :
&nbsp;
http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har 

and detailed information can be found at the link. The aim is to predict how well the participants do some physical activites. __"classe"__ variable stores this information. This is a classification problem.

## Data Exploration 

Import caret library and download the files if it is needed. Read csv files.
```{r}
library(caret)
if(!file.exists("pml-training.csv")){
  download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="pml-training.csv")
}

if(!file.exists("pml-testing.csv")){
  download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="pml-testing.csv")
}  

unPartitionedSet = read.csv(file = "pml-training.csv" ,stringsAsFactors =T,header = T)
testingSet = read.csv(file = "pml-testing.csv", stringsAsFactors =T, header = T)

```

Print sizes of the data sets.
```{r}
dim(unPartitionedSet)
dim(testingSet)
```

Investigate NA values.

```{r}
naControl = colSums(is.na(unPartitionedSet))
unique(naControl)
```

## Data Preprocessing

Columns contain either 0 or 19216 NA values. Columns which contain NA's must be excluded from the data.
Also the first 7 variables are not related to our model. They contain generic knowledge.

```{r}
unPartitionedSet = unPartitionedSet[, -c(1:7)]
unPartitionedSet = unPartitionedSet[, colSums(is.na(unPartitionedSet))==0]
```

Create data partition with 70%.
```{r}
set.seed(111)
trainIndices = createDataPartition(unPartitionedSet$classe,p=0.7)[[1]]
trainingSet = unPartitionedSet[trainIndices,]
validationSet = unPartitionedSet[-trainIndices,]
```

Variables with little variance do not have significant impact.
They should be excluded from training set.

```{r}
lowVar = nearZeroVar(trainingSet)
trainingSet = trainingSet[,-lowVar]
```

Check new dimensions.

```{r}
dim(trainingSet)
```
With data preparation, number of variables in training set is reduced from 160 to 53

&nbsp;

Examine the classe variable. (The project aims to predict it).
```{r}
str(trainingSet$classe)
```

```{r}
table(trainingSet$classe)
```
## Applying Machine Learning Algorithms

Threefold cross validation will be applied.
```{r}
threefold = trainControl(method="cv", number=3, verboseIter=FALSE)
```

### KNN
```{r, cache=TRUE}
set.seed(333)
knnModel = train(classe ~ ., data = trainingSet, method = "knn", trControl=threefold)
predictKnn = predict(knnModel, newdata=validationSet)
confusionMatrix(predictKnn, as.factor(validationSet$classe))
```
This model has 91% accuracy.

### Decision Tree
```{r, cache=TRUE}
set.seed(333)
decisionTreeModel = train(classe ~ ., data=trainingSet, method="ctree", trControl=threefold)
predictDecisionTree = predict(decisionTreeModel, newdata=validationSet)
confusionMatrix(predictDecisionTree, as.factor(validationSet$classe))
```
This model has 91% accuracy too.

### Random Forest

Train random forest with 100 trees.
```{r, cache=TRUE}
set.seed(333)
randomForestModel = train(classe ~ ., data=trainingSet, method="rf", ntree=100 ,trControl=threefold)
predictRandomForest = predict(randomForestModel$finalModel, newdata=validationSet)
confusionMatrix(predictRandomForest, as.factor(validationSet$classe))

```

Random forest has the highest accuracy with 99% . This model will be used for predicting the test set.

## Predicting
```{r}
predict(randomForestModel, newdata = testingSet)

```

The results are predictions on test set. This model is 100% successful at predicting the test set. Note: The test set does not contain __'classe'__  variable to compare, there is a quiz in the MOOC to examine actual values.
